---
output:
  output: rmarkdown::github_document
  #html_document: default
  #pdf_document: default
author: "Samuel Duan"
title: "Red Wine Quality Exploration"
---

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_Library}
library(ggplot2)
library(memisc)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
wqr<-read.csv("wineQualityReds.csv")
```

```{r echo=FALSE}
# the way to create an empty line between paragraphs:
# (space)(space)
# <br>(space)
# (empty line)
```

## Univariate Plots Section  

In this section, some preliminary exploration of the dataset is performed, 
including data set summary and histogram is plotted to show the distribution of 
each particular variable.  
<br>  

#### 0. Summary of the Data Set

```{r echo=FALSE, message=FALSE, warning=FALSE}
str(wqr)
```

This data set contains 1599 observations and 11 chemical attributes that could 
affect the wine quality.    
<br>    

#### 1. Histogram of wine quality
```{r echo=FALSE, Univariate_Plots}
#This idea to color the plots was learned from a post from Oleksii Renov
#on 2/27/2015. http://orenov.github.io/r
ggplot(data=wqr, aes(x=quality))+geom_bar(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$quality)
```
As shown in the plot above that the mode of the wine quality is 5.  
<br>  

#### 2. Histogram of fixed acidity
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=fixed.acidity))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$fixed.acidity)
```
The distribution of fixed acidity is skewed to the right.  
<br>  

#### 3. Histogram of volatile acidity
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=volatile.acidity))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$volatile.acidity)
```
The distribution of volatile acidity is skewed to the right.  
<br>  

#### 4. Histogram of citric acid
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=citric.acid))+geom_histogram(aes(fill=..count..))
ggplot(data=wqr, aes(x=citric.acid))+geom_histogram(aes(fill=..count..))+scale_x_log10()
```
```{r echo=FALSE}
summary(wqr$citric.acid)
```
The distribution of citric acid is skewed to the right, and it seems to be 
multimodal. Thus, it is plotted on a log scale to explore any unseen features.  
<br>  

#### 5. Histogram of residual sugar
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=residual.sugar))+geom_histogram(aes(fill=..count..))
ggplot(data=wqr, aes(x=residual.sugar))+geom_histogram(aes(fill=..count..))+scale_x_log10()
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(wqr$residual.sugar)
```
Since the distriution of residual sugar is significantly skewed to the right and 
concentrated at the left, it is plotted on a log scale to explore explore any 
unseen features.  
<br>  

#### 6. Histogram of chlorides
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=chlorides))+geom_histogram(aes(fill=..count..))
ggplot(data=wqr, aes(x=chlorides))+geom_histogram(aes(fill=..count..))+scale_x_log10()
```
```{r echo=FALSE}
summary(wqr$chlorides)
```
Again, since the distriution of chlorides is significantly skewed to the right, 
it is plotted on a log scale to explore explore any unseen features.  
<br>  

#### 7. Histogram of free sulfur dioxide
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=free.sulfur.dioxide))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$free.sulfur.dioxide)
```

#### 8. Histogram of total sulfur dioxide
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=total.sulfur.dioxide))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$total.sulfur.dioxide)
```
Both the distributions of free and total sulfur dioxide are skewed to the right.  
<br>  

#### 9. Histogram of density
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=density))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$density)
```
While most of the other parameters are positively skewed, the distribution of 
density is quite symmetric.  
<br>  

#### 10. Histogram of pH value
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=pH))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$pH)
```
The distribution of pH value also is symmetric.  
<br>  

#### 11. Histogram of sulphates
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=sulphates))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$sulphates)
```

#### 12. Histogram of alcohol
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=alcohol))+geom_histogram(aes(fill=..count..))
```
```{r echo=FALSE}
summary(wqr$alcohol)
```
Both the distribution of sulphates and alcohol are skewed to the right.  
<br>  

## Univariate Analysis  
### What is the structure of your dataset?  
This data set contains 1599 observations and 11 chemical attributes  that could 
affect the wine quality. All of 11 chemical attributes  are numerical variables.  

Other observations:  
1. The distributions of density and pH value are symmetic while most of the
attributes  are skewed to the right.  
2. Most of the attributes  are unimodal distributions, but Citric acid's has
prominent peaks.  

### What is/are the main feature(s) of interest in your dataset?
The main features of interest after univariate analysis are density and pH
value, since they resemble the distribution of the quality.  

### What other features in the dataset do you think will help support your
investigation into your feature(s) of interest? 
Both the distributions of density and pH value are unimodal.  

### Did you create any new variables from existing variables in the dataset?  
No.  

### Of the features you investigated, were there any unusual distributions? 
Did you perform any operations on the data to tidy, adjust, or change the form
of the data? If so, why did you do this?   
Some of the attributes  (e.g. residual sugar, chlorides) have narrow 
distributions and are positively skewed. Thus, these qualities are plotted on
log scale to explore any hidden features.  

## Bivariate Plots Section  

#### 1. Boxplot: fixed.acidity vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
ggplot(data=wqr, aes(x=quality, y=fixed.acidity, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$fixed.acidity, wqr$quality, summary)
```
No strong correlation is found between fixed acidity and quality.  
<br>  

#### 2. Boxplot: volatile.acidity vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=volatile.acidity, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$volatile.acidity, wqr$quality, summary)
```
Volatile acidity and quality are negatively correlated.  
<br>  

#### 3. Boxplot: citric.acid vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=citric.acid, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$citric.acid, wqr$quality, summary)
```

Citric acid and quality are positively correlated.  
<br>  

#### 4. Boxplot: residual.sugar vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=residual.sugar, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$residual.sugar, wqr$quality, summary)
```
No strong correlation is found between residual sugar and quality.  
<br>  

#### 5. Boxplot: chlorides vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=chlorides, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$chlorides, wqr$quality, summary)
```
No strong correlation is found between residual sugar and quality.  
<br>  

#### 6. Boxplot: free.sulfur.dioxide vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=free.sulfur.dioxide, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$free.sulfur.dioxide, wqr$quality, summary)
```
No strong correlation is found between free sulfur dioxide and quality.  
<br>  

#### 7. Boxplot: total.sulfur.dioxide vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=total.sulfur.dioxide, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$total.sulfur.dioxide, wqr$quality, summary)
```
No strong correlation is found between total sulfur dioxide and quality.  
<br>  

#### 8. Boxplot: density vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=density, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$density, wqr$quality, summary)
```
A week and negative correlation is found between density and quality. Despite
both of the distributions of density and quality are symmetric, it does not
necessarily indicate they are highly correlated.  
<br>  

#### 9. Boxplot: pH vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=pH, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$pH, wqr$quality, summary)
```
Similiar to the previous plot, a week and negative correlation is found between 
density and quality. Their resemblance does not lead to a strong correlation.  
<br>  

#### 10. Boxplot: sulphates vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=sulphates, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$sulphates, wqr$quality, summary)
```
The correlation between Sulphates and quality are positive, yet weak.  
<br>  

#### 11. Boxplot: alcohol vs quality
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqr, aes(x=quality, y=alcohol, fill=quality))+
  geom_boxplot(aes(group=quality))
tapply(wqr$alcohol, wqr$quality, summary)
```
The correlation between alcohol and quality are positive and strong.  
<br>  

#### 12. The correlation plot between chemical attributes  
```{r echo=FALSE, message=FALSE, warning=FALSE}
df=wqr[c("quality", "fixed.acidity", "volatile.acidity", "citric.acid", 
         "residual.sugar", "chlorides", "free.sulfur.dioxide", 
         "total.sulfur.dioxide", "density", "pH", "sulphates","alcohol")]
ggcorr(df, palette = "RdBu", label = TRUE)
```  

The correlation between chemical attributes are calculated to provide a 
quantitative insight on their relationship. Both this chart and the above 
boxplots draw the same conclusion: volatile acidity and alcohol show
stronger correlation with the wine quality. While the correlation between
sulphate and wine quality is 0.3, many outliers were found in the boxplot.  
<br>  

## Bivariate Analysis  

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?   
1. When we explored the dataset by univariate plots, density and pH value are 
the only two attributes show symmetric distributions. At that time, it was 
suspected that they may have a stronger correlation with wine quality. 
2. However, as discussed briefly in the section of bavariate plots, volatile 
acidity and alcohol, rather than density and pH value show stronger correlation 
with the wine quality. 
3. Thus, the features of interest now are the votalie acidity and alcohol.  
  

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?  
3 chemical attributes show strong correlation with fixed acidity (which is 
stronger than those with wine quality) and they are pH value, density and 
citric acid.  

### What was the strongest relationship you found?
Among the chemical attribute: (1) pH value and fixed acidity, (2) density and 
fixed acidity, (3) citric acid and fixed acidity  
Between the wine quality and a sole chemical attribute: alcohol and wine quality  
<br>

## Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
wqr$newquality = ifelse(wqr$quality>5, 1, 0)
ggplot(wqr, aes(x=pH, y=fixed.acidity, group=newquality, colour=newquality)) + geom_point()
```
  
In this plot, I create a new variable to indicate the wine quality is high or
low. If wine quality is larger than 5, 1 is assigned, indicating high quality. 
If it is smaller than 5, 0 is assigned, indicating low quality. Then 
fixed acidity is plotted against pH value for both high and low quliaty. As 
shown in the plot, wine quality is in effect independent of the correlation 
between fixed acidity and pH value. Fixed acidity and pH value has a significant
correlation as shown in the previous discussion.  
  
Likewise, I continue to go through the other two sets of attributes who have stronger correlations and see if they are corelcated with wine quality or not.  

What if I plot the pH value vs. fixed acidity by every value of wine quality?
```{r echo=FALSE}
#This idea to employ facet_wrap was learned from a post from Oleksii Renov
#on 2/27/2015. http://orenov.github.io/r
ggplot(aes(x=pH, y=fixed.acidity, color = quality), data = wqr) + geom_point() + 
  facet_wrap(~quality) + geom_smooth(colour='black')
```
<br>  
Still, no particular pattern can be found.  
  
Let's go through the other two sets of correlated attributes to see if there
is any visible patterns.
  
```{r echo=FALSE}
ggplot(wqr, aes(x=density, y=fixed.acidity, group=newquality, 
                colour=newquality)) + geom_point()
```
```{r echo=FALSE}
ggplot(wqr, aes(x=citric.acid, y=fixed.acidity, group=newquality, 
                colour=newquality)) + geom_point()
```
No particular pattern can be found.  

<br>  
  
Now I employ stepwise model selection to determine which chemical attributes
have an important role on the wine quality.  
```{r echo=FALSE}
library(MASS)
fit <- lm(quality~fixed.acidity+volatile.acidity+citric.acid+residual.sugar
+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates
+alcohol, data=wqr)

stepfit <- step(fit, direction="both")
stepfit$anova # display results
plot(fit$residuals)
```
  
According the result from the model selection, the amount of (1) volatile 
acidity, (2) chlorides, (3) free.sulfur.dioxide, (4) total.sulfur.dioxide, 
(5) pH, (6) sulphates, and (7) alcohol would have an affect on the wine quality.  

In the prior bivariate analysis, alcohol is the only attribute that shows strong
correlation with the wine quality. At first, it seems to be contray to the model
selection result. As I look into the model selection as shown above, the removal
of alcohol would cause a huge decrease in AIC value, while the removal of others
affect AIC less significantly. In fact, this list shows roughly the influence of
that attribute on the wine quality in an ascending order (that is from less 
important to more important).  
<br>  
  
## Multivariate Analysis  
  
### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
In the prior discussion, alcohol influences the wine quality the most. This is
confirmed by both the correlation plot in the bivariate analysis and the 
regression result in this section.  

### Were there any interesting or surprising interactions between features?
I am quite surprise that alcohol influences the wine quality so greatly,
while the other attributes do not show the same degree of effect.

------

## Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data=wqr, aes(x=quality, y=alcohol, fill=quality))+
  geom_boxplot(aes(group=quality)) + 
  labs(title="Alcohol by quality", x="Quality", y = "Alcohol (% by volume)") +
  theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        plot.title=element_text(size=18),
        legend.text=element_text(size=11))
```

### Description One
A boxplot is drawn to show the relationship between the amount of alcohol 
(% by volume) and the wine quality. It is visible that these two are correlated
positively. Colored filling is employed to increase the readibility.  

### Plot Two
```{r echo=FALSE, Plot_Two, fig.width = 12}
df=wqr[c("quality", "fixed.acidity", "volatile.acidity", "citric.acid", 
         "residual.sugar", "chlorides", "free.sulfur.dioxide", 
         "total.sulfur.dioxide", "density", "pH", "sulphates","alcohol")]
ggcorr(df, palette = "RdBu", label = TRUE, hjust = 0.8)+
  labs(title="Corrgram for wine quality")+  
  theme(plot.title=element_text(size=18))
```

### Description Two
A correlation matrix plot is drawn, showing the correlation between all the 
attributes. Particularly, the correlation with the wine quality is of our
interest. Alcohol, volatile acidity, and sulphates are the top three factors
having significant relation with the wine quality.  

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x=pH, y=fixed.acidity, color = quality), data = wqr) + geom_point() + 
  facet_wrap(~quality) + geom_smooth(colour='black') + 
  ylab("Fixed acidicity tartaric acid - g / dm^3") +
  xlab("pH") + ggtitle("Fixed acidicity vs pH by Quality") + 
  theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        plot.title=element_text(size=18),
        legend.text=element_text(size=11))
```

### Description Three
This is to explore two high correlation variables (Fixed acidity and pH) by 
quality. However, we cannot observe unsual things. Although the plots at low
quality do not have sufficient data points, overall we can see some stability
between these two high correlation variables. Lastly, through this project,
I realized coloration is of great importance. Coloring plots would make 
everything logically flow together.
------

## Reflection
This red wine quality was created using red wine samples (1599 observations as 
total), based on the median of at least 3 evaluation made by wine experts. 11 
quantitative wine attributes, which were generated through more objective tests,
are included in this data set. As I explored the data set and tried to find out
the correlation between 11 attributes and wine quality through various plots and
testing, the amount of alcohol content in the wine is the most influencial factor
in determing the wine quality. A predictive model is also developed, 7 
variables are included to give the best fitting result, while alcohol remains
the most critical among them.
